Secure Composition of Robust and Optimising Compilers

### 2023-06-17 popl/main.tex
Baseline def

\usepackage[colorinlistoftodos]{todonotes}
\usepackage[inference]{semantic}
\usepackage{fontawesome5}
\usepackage{glossaries}
\usepackage{cleveref}
\usepackage{stmaryrd}
\usepackage{listings}
\usepackage{xspace}
\usepackage{tikz}
\usepackage{bm}

@@

This section defines several secure compilers, each of which robustly preserves a different property of interest (\Cref{fig:pipeline}).
The power of the framework (\Cref{sec:compprop,sec:compcomp}) is demonstrated by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
Concretely, the section starts with a base-line definition for the used languages and extends that to define languages $\src{L_{TMS}}$ and $\trg{L}$.
For programmers using $\src{L_{TMS}}$, their mental model should be that their code is temporal memory safe and cryptographic constant time.
The former is proven to be enforced by the static semantics of $\src{L_{TMS}}$, where each well-typed program is temporal memory safe.
The latter is completely untyped.
Having languages defined, this section continues with a compiler $\cc{\src{L_{\tmssafe}}}{\trg{L}}$ and proves that $\cc{\src{L_{\tmssafe}}}{\trg{L}}$ robustly preserves \gls{tms} (\Cref{subsec:cs:tms}).
For readability purposes, it defines $\irl{L_{MS}}$ to be exactly the same as $\trg{L}$ and continues to present a compiler $\cc{\trg{L}}{\irl{L_{MS}}}$ that robustly preserves \gls{sms} (\Cref{subsec:cs:ms}).
Now leaveraging the power of the framework that this paper introduces (\Cref{sec:compcomp,sec:compprop}), a compiler $\rtp{\cc{\src{L_{TMS}}}{\irl{L_{MS}}}}{\mssafe}$ is attained.
The next steps in the compilation pipeline are two secure compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ that perform \glsfirst{dce} and \glsfirst{cf}, respectively, to model an optimzing pipeline happening somewhere in the compilation chain (\Cref{subsec:cs:opts}).
Finally, this section defines the language $\obj{L_{\scctsafe}}$ that is, in contrast to the other languages, not cryptographic constant time by definition.
Concluding, a compiler $\cc{\irl{L_{MS}}}{\obj{L_{CCT}}}$ that robustly preserves \gls{cct} (\Cref{subsec:cs:scct}) is shown and composed with $\cc{\src{L_{TMS}}}{\irl{L_{MS}}}$ and the optimizing compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$.
In turn, this yields $\cc{\src{L_{TMS}}}{\obj{L_{CCT}}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.

>>>>

\usepackage[colorinlistoftodos]{todonotes}
\usepackage[inference]{semantic}
\usepackage{fontawesome5}
\usepackage{listofitems}
\usepackage{glossaries}
\usepackage{cleveref}
\usepackage{stmaryrd}
\usepackage{marvosym}
\usepackage{listings}
\usepackage{xspace}
\usepackage{xfrac}
\usepackage{tikz}
\usepackage{bm}

@@

This section defines several secure compilers, each of which robustly preserves a different property of interest (\Cref{fig:pipeline}).
The power of the framework (\Cref{sec:compprop,sec:compcomp}) is demonstrated by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
Concretely, the section starts with a base-line definition for the used languages (\Cref{subsec:cs:defs}) and extends that to define languages $\src{L_{TMS}}$ and $\trg{L}$.
For programmers using $\src{L_{TMS}}$, their mental model should be that their code is temporal memory safe and cryptographic constant time.
The former is proven to be enforced by the static semantics of $\src{L_{TMS}}$, where each well-typed program is temporal memory safe.
The latter is completely untyped.
Having languages defined, this section continues with a compiler $\cc{\src{L_{\tmssafe}}}{\trg{L}}$ and proves that $\cc{\src{L_{\tmssafe}}}{\trg{L}}$ robustly preserves \gls{tms} (\Cref{subsec:cs:tms}).
For readability purposes, it defines $\irl{L_{MS}}$ to be exactly the same as $\trg{L}$ and continues to present a compiler $\cc{\trg{L}}{\irl{L_{MS}}}$ that robustly preserves \gls{sms} (\Cref{subsec:cs:ms}).
Now leaveraging the power of the framework that this paper introduces (\Cref{sec:compcomp,sec:compprop}), a compiler $\rtp{\cc{\src{L_{TMS}}}{\irl{L_{MS}}}}{\mssafe}$ is attained.
The next steps in the compilation pipeline are two secure compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ that perform \glsfirst{dce} and \glsfirst{cf}, respectively, to model an optimzing pipeline happening somewhere in the compilation chain (\Cref{subsec:cs:opts}).
Finally, this section defines the language $\obj{L_{\scctsafe}}$ that is, in contrast to the other languages, not cryptographic constant time by definition.
Concluding, a compiler $\cc{\irl{L_{MS}}}{\obj{L_{CCT}}}$ that robustly preserves \gls{cct} (\Cref{subsec:cs:scct}) is shown and composed with $\cc{\src{L_{TMS}}}{\irl{L_{MS}}}$ and the optimizing compilers $\cc[_{\gls{dce}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$ and $\cc[_{\gls{cf}}]{\irl{L_{MS}}}{\irl{L_{MS}}}$.
In turn, this yields $\cc{\src{L_{TMS}}}{\obj{L_{CCT}}}$, a compiler that robustly preserves a combination of \gls{ms} and \gls{cct}.

\subsection{Preliminary Definitions}\label{subsec:cs:defs}
\begin{gather*}
  \begin{aligned}
  \mi{(Finals)}~\finalexprnoerr&\bnfdef\ \valueexpr \mid x\hspace{0.5cm}
  \mi{(Results)}~\finalexpr\bnfdef\ \finalexprnoerr \mid stuck\hspace{0.5cm}
  \mi{(Values)}~\valueexpr\bnfdef\ n\\
  \mi{(Expressions)}~\expr&\bnfdef\ \finalexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{x}{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \\
    & \text{where}\ \lbinop{}{}\in\left\{+,-,\times,<,/\right\}\\
  \mi{(Functions)}~\function&\bnfdef\ \lfunction{g}{x}{e}\hspace{0.5cm}
  \mi{(Types)}~\type\bnfdef\ \natt\hspace{0.5cm}
  \mi{(Locations)}~\loc\in\nat \\
  \mi{(Eval.~Ctx.)}~\ectx&\bnfdef\ \hole{\cdot} \mid \lbinop{\ectx}{\expr} \mid \lbinop{\valueexpr}{\ectx} \mid \lget{x}{\ectx} \mid \lnew{x}{\ectx}{\expr} \mid \lset{x}{\ectx}{\expr}\\
    & \mid \lset{x}{\valueexpr}{\ectx} \mid \lreturn{\ectx} \mid \lcall{g}{\ectx} \mid \lifz{\ectx}{\expr[_{1}]}{\expr[_{2}]} \\
  \mi{(Poison)}~\poison&\bnfdef\ \poisoned \mid \poisonless \hspace{0.5cm}
  \mi{(Store)}~\store\bnfdef\ \hole{\cdot} \mid \storeel{x}{\loc}{\sandboxtag}{\poison}{n},\store \\
  \mi{(Transfer~Tag)}~\comm&\bnfdef\ \nocomm \mid\ \ctxtocomp \mid\ \comptoctx \hspace{0.5cm}
  \mi{(Heaps)}~\heap\bnfdef\ \hole{\cdot} \mid n,\heap \hspace{0.5cm}
  \mi{(Library)}~\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
  \mi{(Component~Names)}~\commlib&\bnfdef\ \hole{\cdot} \mid g,\commlib \hspace{0.5cm}
  \mi{(Continuation~Stack)}~\ectxstack\bnfdef\ \hole{\cdot} \mid \tup{\ectx}{g},\ectxstack \\
  \mi{(Flow~State)}~\cfstate&\bnfdef\ \tup{\commlib}{\library,\ectxstack} \hspace{0.5cm}
  \mi{(Memory~State)}~\memstate\bnfdef\ \tup{\heap[^{\ctx}]}{\heap[^{\comp}],\store} \hspace{0.5cm}
  \end{aligned}
\end{gather*}
